Spitzer, Gaia and the Potential of the Milky Way

### 2013-07-26 text/paper1/paper.tex
not entirely sure..

This approach was originally propose by \citep{jzsh1999} and designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$. Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris. Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure; and the test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.

@@

\subsection{The algorithm: Rewinder}
Quantifying this method requires a sample of stars with known full
space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements of all position and velocity components for these stars \emph{today} at $t=0$), the orbital parameters
for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
functional form for the potential, $\Phi({\boldsymbol\theta})$. For a
given set of potential parameters, $\boldsymbol\theta$, the orbits of the stars and
progenitor are integrated backwards for several Gigayears (determined from an
estimate of the interaction time of the system). The stars and
progenitor system are treated as test particles, so that the debris
stars do not become physically re-bound to the satellite. Instead, at each
timestep $t_j$, for each particle $i$, a set of normalized, relative
phase-space coordinates are
computed 
\begin{equation}
  \bs{q}_{i} = \frac{\bs{x}_{i} -
    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,;\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
    \bs{v}_{p}}{v_{\rm esc}}
\end{equation}
where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the phase-space coordinates for the particles and progenitor, respectively. These definitions require an estimate of the mass of the
satellite, $m_{sat}$, which, combined with the orbital radius of the satellite, $R$, and the computed enclosed mass
of the potential within $R$, $m_{enc}$, sets the instantaneous tidal radius and
escape velocity,
\begin{equation}
  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
  v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
\end{equation}
Qualitatively, when the distance in this normalized six-dimensional space,
$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
likely recaptured by the satellite. JZSH imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization. What follows is a description of an updated procedure with a more natural choice for an objective function [TODO...].

As a statistical measure of how many stars have recombined, we
use the minimum phase-space vectors --- for
each star over the individual, integrated orbits --- as samples from an underlying six-dimensional distribution. For each star, $i$,
the phase-space distance, $D_{\rm ps}$, is computed at each timestep
$t_{j}$, and the vector with the minimum phase-space distance is stored
\begin{align}
  t^*_{i} &= \argmin_{t} D_{{\rm ps},i}\\
  \bs{A}_{i} &= (\bs{q}_{i}(t^*_{i}),\bs{p}_{i}(t^*_{i})).
\end{align}
Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
where $k\in[1,6]$. Intuitively, the variance of this distribution of minimum distance vectors
will be larger for orbits integrated in an incorrect potential
relative to the distribution computed from the `true' orbital history
of the stars; in an incorrect potential, small differences in the
orbits of the stars relative to the orbit of the progenitor cause them
to spread out in this relative, normalized phase space. Thus, the
\emph{generalized variance} of the distribution, computed for a given
set of potential parameters, $\bs{\theta}$, is a natural choice for
the scalar objective function used in constraining the potential of
the Milky Way,
\begin{equation}
  f(\bs{\theta})=\det(\mathrm{Cov}( \bs{A}_{ik})).
\end{equation}

>>>>

This approach was originally proposed by \citep{jzsh1999} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$. Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris. Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure. The test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.

@@

\subsection{The algorithm: Rewinder}
Quantifying this method requires a sample of stars with known full
space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements of all position and velocity components for these stars \emph{today} at $t=0$), the orbital parameters
for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
functional form for the potential, $\Phi({\boldsymbol\theta})$. For a
given set of potential parameters, $\boldsymbol\theta$, the orbits of the stars and
progenitor are integrated backwards for several Gigayears (determined from an
estimate of the interaction time of the system). The stars and
progenitor system are treated as test particles, so that the debris
stars do not become physically re-bound to the satellite. Instead, at each
timestep $t_j$, for each particle $i$, a set of normalized, relative
phase-space coordinates are
computed 
\begin{equation}
  \bs{q}_{i} = \frac{\bs{x}_{i} -
    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,;\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
    \bs{v}_{p}}{v_{\rm esc}}
\end{equation}
where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the phase-space coordinates for the particles and progenitor, respectively. These definitions require an estimate of the mass of the
satellite, $m_{sat}$, which, combined with the orbital radius of the satellite, $R$, and the computed enclosed mass
of the potential within $R$, $M_{enc}$, sets the instantaneous tidal radius and
escape velocity,
\begin{equation}
  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3M_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
  v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
\end{equation}
Qualitatively, when the distance in this normalized six-dimensional space,
$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
likely recaptured by the satellite. JZSH imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization. What follows is a description of an updated procedure with a more natural choice for an objective function [TODO...].

As a statistical measure of how many stars have recombined, we
treat the minimum phase-space vectors (for
each star over the individual, integrated orbits) as samples from an underlying six-dimensional distribution. For each star, $i$,
the phase-space distance, $D_{\rm ps}$, is computed at each timestep
$t_{j}$, and the vector with the minimum phase-space distance is stored
\begin{align}
  t^*_{i} &= \argmin_{t} D_{{\rm ps},i}\\
  \bs{A}_{i} &= (\bs{q}_{i}(t^*_{i}),\bs{p}_{i}(t^*_{i})).
\end{align}
Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
where $k\in[1,6]$. Intuitively, the variance of this distribution
will be larger for orbits integrated in an incorrect potential
relative to the distribution computed from the `true' orbital history
of the stars: in an incorrect potential, small differences in the
orbits of the stars relative to the orbit of the progenitor cause them
to spread out in this relative, normalized phase space. Thus, the
\emph{generalized variance} of the distribution, computed for a given
set of potential parameters, $\bs{\theta}$, is a natural choice for
the scalar objective function used in constraining the potential of
the Milky Way,
\begin{equation}
  f(\bs{\theta})=\det(\mathrm{Cov}( \bs{A}_{ik})).
\end{equation}
